### 2016-05-06 presentation.tex
add presentation, copied from SOP ES_NASM_06_16_13

%\documentclass{beamer} 
\documentclass[handout]{beamer} 
\usetheme{Ilmenau}
\usepackage{graphicx,verbatim,hyperref}
\usepackage{textpos}

\usecolortheme{beaver}
\useinnertheme{default}
\setbeamertemplate{itemize item}[triangle]
\setbeamertemplate{itemize subitem}[triangle]
\setbeamertemplate{itemize subsubitem}[circle]
\setbeamertemplate{enumerate items}[default]
\setbeamertemplate{blocks}[upper=block head,rounded]
\setbeamercolor{item}{fg=black}
\usefonttheme{serif} %should allow ccfonts to take effect

\usepackage{cite}
\usepackage{times, verbatim,xcolor,bm}
\usepackage{amsbsy,amssymb, amsmath, amsthm}
\usepackage{booktabs}
%David miller's fonts
	\usepackage[T1]{fontenc}
	\usepackage[boldsans]{ccfonts}
	\usepackage[euler-hat-accent]{eulervm}

\newcommand{\al}{\alpha}
\newcommand{\expect}{\mathbb{E}}
\newcommand{\Bt}{B(\bm{\tau^a})}
\newcommand{\bta}{\bm{\tau^a}}
\newcommand{\btn}{\bm{\tau^{tw}}}
\newcommand{\ga}{\gamma}
\newcommand{\ve}{\varepsilon}
\newcommand{\ta}{\theta}

\newenvironment{changemargin}[2]{% 
  \begin{list}{}{% 
    \setlength{\topsep}{0pt}% 
    \setlength{\leftmargin}{#1}% 
    \setlength{\rightmargin}{#2}% 
    \setlength{\listparindent}{\parindent}% 
    \setlength{\itemindent}{\parindent}% 
    \setlength{\parsep}{\parskip}% 
  }% 
  \item[]}{\end{list}}

	\let\Tiny=\tiny

\title[Temporary Trade Barriers: When Will They End?\hspace{2.15in}\insertframenumber/\inserttotalframenumber]{Temporary Trade Barriers: \\ When Will They End?}
\author[Kristy Buzard]{\texorpdfstring{Kristy Buzard\newline Syracuse University and The Wallis Institute  \newline\url{kbuzard@syr.edu}}{Kristy Buzard}}
\date{May 13, 2016}

\begin{document}
\maketitle
%\insertpresentationendpage removed b/c of appendix

\section{Overview}
\subsection{Preview}
\begin{frame}
\frametitle{Political Economy of Trade Agreements}
In most models, a single actor in each country makes trade policy. Modeling policy-making power more richly:
\pause
\begin{itemize}[<+->]
	\item reconciles `Protection for Sale' model to the data
  \item generates lobbying for the purpose of disrupting trade agreements  
	\item adds new role for trade agreements 
\end{itemize}
\end{frame}

\begin{frame}{Preview of Results}
\begin{itemize}[<+->]
	\item Higher trade agreement tariffs induce lobbies to decrease effort 
		\begin{itemize}
			\item Reduces probability of trade war / ratification failure
		\end{itemize}
	\item Executives negotiate tariffs to discourage lobbying
		\begin{itemize}
			\item Produces higher tariffs than executives prefer
			\item Reduces equilibrium lobbying effort
			\item Trade agreement as political commitment device
		\end{itemize}
	\item Tariff levels and optimal lobbying depend on how much political uncertainty is present
\end{itemize}
\end{frame}

\begin{frame}{Outline of Talk}
\begin{enumerate}[<+->]
	\item Background and Motivation
	\item Economic and Political Structure
	\item Results for a Special Case: Certainty
	\item Full Results: Political Uncertainty
	\item Conclusion
\end{enumerate}
\end{frame}

\begin{frame}{Standard Unitary Government Model}
Grossman $\&$ Helpman government welfare function:
				\[
				  C + aW
				\]
\pause
		\begin{itemize}[<+->]
			\item Government auctions tariffs/subsidies: lobbies pay for $a$-weighted welfare loss
			\item Estimated $a$'s far too large
				\begin{itemize}
					\item Implies that governments respond very little to lobbying effort
					\item Inconsistent with observed large tariffs relative to small lobbying effort
				\end{itemize}
		\end{itemize}
\end{frame}

\begin{frame}{Trade Agreements and Separation of Powers}
  \pause
  Take special case of `Protection for Sale' model (single lobby) and generalize government structure
  \pause
		\begin{itemize}[<+->]
			\item Executives can make trade agreement
			\item Legislatures retain final authority: can break agreement and start trade war / fail to ratify
			\item Trade war is subgame of the trade-agreement-setting game
				\begin{itemize}
  				\item Executives consider lobbying incentives and possibility of trade war when forming agreement
				\end{itemize}
		\end{itemize}
\end{frame}

\subsection{}
\begin{frame}
\frametitle{Related Literature}
\small Protection for Sale: Grossman $\&$ Helpman (1994)
\begin{itemize}
  \item \footnotesize Empirics: Goldberg $\&$ Maggi (1999), Gawande $\&$ Bandyopadhyay (2000), Mitra, Thomakos, $\&$  Ulubasoglu (2002)
  \item \footnotesize Mitra, Thomakos, $\&$  Ulubasoglu (2006), Bombardini (2008)
	\item \footnotesize Trade Wars and Trade Talks: Grossman $\&$ Helpman (1995)
\end{itemize}

\vskip.05in
\small Political economy shocks
\begin{itemize}
	\item \footnotesize Feenstra $\&$ Lewis (1991), Bagwell $\&$ Staiger (2001, 2005)
\end{itemize}

\vskip.05in
\small Separated powers
\begin{itemize}
	\item \footnotesize Mansfield, Milner $\&$ Rosendorff (2000), Song (2008)
\end{itemize}

\vskip.05in
\small Political uncertainty
\begin{itemize}
	\item \footnotesize Milner $\&$ Rosendorff (1997), Le Breton $\&$ Zaporozhets (2007) %legislators are agents, symmetric case very hard
\end{itemize}

		%\item Autocracy vs. Democracy: Mitra, Thomakos and Ulubasoglu (2002), Aidt and Gassebner (2010)
\end{frame}

\section{Model}
\subsection{Economic and Political Structure}
\begin{frame}{Timeline}
\pause
\begin{enumerate}
	\item {\bfseries Formation}
		\begin{enumerate}[i.]
			\pause
			\item Executives set trade policy in international agreement
		\end{enumerate}
	\pause
	\item \textbf{Ratification / Maintenance}
		\begin{enumerate}[i.]
			\pause
			\item Firms lobby legislatures to break agreement
			\pause
			\item Uncertainty is resolved
			\pause
			\item Legislatures decide to break or abide by agreement
			\pause
		\end{enumerate}
	\item \textbf{Trade War} (if agreement is broken)
	\pause
		\begin{enumerate}[i.]
			\item Firms lobby legislatures to set high trade-war tariff
			\pause
			\item Uncertainty is resolved
			\pause
			\item Legislatures decide trade-war tariff
		\end{enumerate}
\pause
	\item Private actors make production, consumption decisions
\end{enumerate}
\end{frame}

\begin{frame}{Economy}
\begin{itemize}
	\item Two countries: home and foreign (${}^*$)
	\item Separable in three goods: $X$ and $Y$ (traded) and numeraire
	\item Demand identical for both goods in both countries
	\item Supply: $Q_X^*(P_X) > Q_X(P_X)$ $\forall P_X$; symmetric for $Y$ 
		\begin{itemize}
			\item Home net importer of $X$, net exporter of $Y$
		\end{itemize}
\end{itemize}

\vskip.2in
\pause
Home levies $\tau$ on $X$, Foreign levies $\tau^*$ on $Y$
\pause
\begin{itemize}
	\item $P_X=P_X^W + \tau$ and $\pi_X(P_X)$ increasing in $\tau$
\end{itemize}

\pause
\vskip.2in
Non-tradable specific factors motivates political activity

\end{frame}

\begin{frame}{Political Structure}
In each country (focus on Home):
\pause
\begin{itemize}[<+->]
	\item A Unitary Executive
		\begin{itemize}
			\item Delegated authority to make trade agreement
		\end{itemize}
		%\pause
	\item A Non-unitary Legislature
		\begin{itemize}[<+->]
			\item Can withdraw delegation, break agreement, and set trade-war tariff
			\item Susceptible to influence of lobbying
			\item Decision determined by median legislator
		\end{itemize}
%\pause
	\item A Single Lobby
		\begin{itemize}
			\item Represents import-competing sector, $X$ ($Y$ in foreign)
		\end{itemize}	
\end{itemize}

\end{frame}

\subsection{The Players}
\begin{frame}
\frametitle{Executive Branch}
Trade agreement negotiated by unitary executive:
$$W_E = \mathit{CS}_X(\tau) + \mathit{CS}_Y(\tau^*) + \ga_E \pi_X(\tau) + \pi_Y(\tau^*) + \mathit{TR}(\tau)$$

\vskip.2in
\begin{itemize}
	\item $\mathit{CS_i(\cdot)}$: consumer surplus
	\item $\pi_X(\tau)$: profits of import-competing industry
	\item $\pi_Y(\tau^*)$: profits of exporting industry
	\item $\mathit{TR}(\tau)$: tariff revenue
	\item $\ga_E$: weight on profits in the import-competing industry
\end{itemize}

\end{frame}

\begin{frame}{Non-Unitary Legislature}
  Decisions determined by preferences of Median Legislator:
\[
  W_{\mathit{ML}} = \mathit{CS}_X(\tau) + \mathit{CS}_Y(\tau^*) + \ga(e,\ta) \pi_X(\tau) + \pi_Y(\tau^*) + \mathit{TR}(\tau)
\]
\vskip-.1in
\pause
\begin{itemize}
	\item $\ga(e,\ta)$: weight on import-competing industry profits
		\begin{itemize}
			\pause
			\item $e$: lobbying effort
			\pause
			\item $\ta$: uncertain element in determination of ML's identity
		\end{itemize}
\end{itemize}

\pause
\vskip.1in
\begin{beamerboxesrounded}[upper=palette tertiary, shadow=true]{Assumptions on $\ga(e,\ta)$}
\begin{enumerate}
  \pause
	\item $\ga(e,\ta)$ is increasing and concave in $e$ for all $\ta \in \Theta$.
  \pause
  \item $\ga(e,\ta) \geq \ga_E \geq 1 \ \forall \ta$ 
\end{enumerate}
\end{beamerboxesrounded}
\end{frame}

\begin{frame}
\frametitle{Lobby}
\vskip.2in
Lobby chooses effort to maximize:
\begin{multline*}
  \left\{1-\Pr\left[ \text{TradeWar}(e_b,\bta)\right]\right\} \ \pi(\tau^a) \\
     + \Pr\left[ \text{TradeWar}(e_b,\bta) \right] \left[ \pi(\tau^{\mathit{tw}}) - e_{tw} \right] - e_b
\end{multline*}

\vskip.2in
\begin{itemize}
	\item $e_b$: Lobbying effort for break decision
	\item $e_{tw}$: Lobbying effort for trade-war tariffs
	\item $\bm{\tau^a} = (\tau^a,\tau^{*a})$: tariff under an agreement
	\item $\bm{\tau^{\mathit{tw}}} = (\tau^{\mathit{tw}},\tau^{*\mathit{tw}})$: tariff when agreement is broken
\end{itemize}

\end{frame}

\section{Certainty}
\subsection{Results: The Certain Case}
\begin{frame}{Timeline}
\begin{enumerate}
	\item \textbf{Formation}
		\begin{enumerate}[i.]
			\item Executives set trade policy in international agreement
		\end{enumerate}
	\item \textbf{Ratification / Maintenance}
		\begin{enumerate}[i.]
			\item Firms lobby legislatures to break agreement
			\item Legislatures decide to break or abide by agreement
		\end{enumerate}
	\item {\color{gray} \textbf{Trade War} (if agreement is broken)}
		\begin{enumerate}[i.]
			\item {\color{gray} Firms lobby legislatures to set high trade-war tariff}
			\item {\color{gray} Legislatures decide trade-war tariff}
		\end{enumerate}
	\item {\color{gray} Private actors make production, consumption decisions}
\end{enumerate}
\end{frame}

\begin{frame}{Highlight: Separation of Powers}
\pause
\textbf{Legislature}
\pause
\begin{itemize}
	\item Breaks agreement if median legislator prefers $\bm{\tau^{\mathit{tw}}}$ to $\bta$
\end{itemize}

\pause
\vskip.1in
\textbf{Lobby}
\pause
\begin{itemize}[<+->]
	\item Given the $\bta$ it faces, lobby knows what $e_b$ is required to break the agreement
 	\item Lobby pays this $e_b$ if: \hskip.2in $\bm{\pi}(\bm{\tau^{\mathit{tw}}}) \bm{- e_{tw}  - e_b > \pi}(\bm{\tau^a})$
\end{itemize}

\pause
\vskip.13in
\textbf{Executives}
\pause
\begin{itemize}[<+->]
	\item Set $\bta$ to make paying $e_b$ unprofitable
		\begin{itemize}
			\item[$\Rightarrow$] $e_b=0$, agreement remains in force
		\end{itemize}
	\item High tariffs, no lobbying, no trade disruptions
\end{itemize}
\end{frame}

\section{Full Results}
\begin{frame}{Political Uncertainty Illustration}
\pause
An Example (Bagwell $\&$ Staiger 2005)
\pause
\begin{itemize}[<+->]
	\item $D(P_i) = 1 - P_i$
	\item $Q_X(P_X) = \frac{P_X}{2}$, $Q_Y(P_Y) = P_Y$
	\item $P_X^W = \frac{4-3\tau}{7}$, $P_X = \frac{4+4\tau}{7}$
	\item $\ga(e,\ta) = 1.25 + e^{0.2} + \ta$ 
		\begin{itemize}
			\item $\ta \sim U[-0.25,0.25]$
		\end{itemize}
	\item $\ga_E = 1$
\end{itemize}

\end{frame}

\subsection{Ratification / Maintenance Stage}
\begin{frame}{Timeline}
\begin{enumerate}
	\item {\color{gray} \textbf{Formation}}
		\begin{enumerate}[i.]
			\item {\color{gray} Executives set trade policy in international agreement}
		\end{enumerate}
	\item \textbf{Ratification / Maintenance}
		\begin{enumerate}[i.]
			\item {\color{gray} Firms lobby legislatures to break agreement}
			\item Legislatures decide to break or abide by agreement
		\end{enumerate}
	\item {\color{gray} \textbf{Trade War} (if agreement is broken)}
		\begin{enumerate}[i.]
			\item {\color{gray} Firms lobby legislatures to set high trade-war tariff}
			\item {\color{gray} Legislatures decide trade-war tariff}
		\end{enumerate}
	\item {\color{gray} Private actors make production, consumption decisions}
\end{enumerate}
\end{frame}

\begin{frame}{Legislature}
  Legislature breaks trade agreement if median legislator's utility is higher under trade war than trade agreement
  \begin{itemize}
		\item Median legislator's identity is uncertain through $\ta$
	\end{itemize}

	\pause
  \vskip.2in
  Probability that Legislature breaks agreement: \\
	\pause
  \vskip.1in
    $b(e_b,\bta,\btn,\ta)$ probability median legislator prefers $\tau^{tw}$ to $\tau^a$ for a given $\ta$
\end{frame}

\begin{frame}{Timeline}
\begin{enumerate}
	\item {\color{gray} \textbf{Formation}}
		\begin{enumerate}[i.]
			\item {\color{gray} Executives set trade policy in international agreement}
		\end{enumerate}
	\item \textbf{Ratification / Maintenance}
		\begin{enumerate}[i.]
			\item Firms lobby legislatures to break agreement
			\item {\color{gray} Legislatures decide to break or abide by agreement}
		\end{enumerate}
	\item {\color{gray} \textbf{Trade War} (if agreement is broken)}
		\begin{enumerate}[i.]
			\item {\color{gray} Firms lobby legislatures to set high trade-war tariff}
			\item {\color{gray} Legislatures decide trade-war tariff}
		\end{enumerate}
	\item {\color{gray} Private actors make production, consumption decisions}
\end{enumerate}
\end{frame}

\subsection{Formation Stage}
\begin{frame}{Timeline}
\begin{enumerate}
	\item \textbf{Formation}
		\begin{enumerate}[i.]
			\item Executives set trade policy in international agreement
		\end{enumerate}
	\item {\color{gray} \textbf{Ratification / Maintenance}}
		\begin{enumerate}[i.]
			\item {\color{gray} Firms lobby legislatures to break agreement}
			\item {\color{gray} Legislatures decide to break or abide by agreement}
		\end{enumerate}
	\item {\color{gray} \textbf{Trade War} (if agreement is broken)}
		\begin{enumerate}[i.]
			\item {\color{gray} Firms lobby legislatures to set high trade-war tariff}
			\item {\color{gray} Legislatures decide trade-war tariff}
		\end{enumerate}
	\item {\color{gray} Private actors make production, consumption decisions}
\end{enumerate}
\end{frame}

\begin{frame}{Total Break Probability}
$B(\bta) = b(e_b(\bta),\bta)$
\pause
\begin{itemize}[<+->]
	\item Both direct and indirect effects of raising $\bta$ are negative
		\begin{itemize}
			\item $B(\bta)$ decreasing in $\bta$
		\end{itemize}
\end{itemize}
\pause
\vskip.2in
\begin{beamerboxesrounded}[upper=palette tertiary, shadow=true]{Result 3}
   The total probability that the trade agreement will be broken is decreasing in the trade agreement tariffs.
\end{beamerboxesrounded}

\end{frame}

\begin{frame}{Executives' Tradeoff}
\[
  \Pr\left[ \text{TradeWar}(\bta) \right] \bm{W_E}(\btn) + \left\{1- \Pr\left[ \text{TradeWar}(\bta) \right]\right\} \bm{W_E}(\bta)
\]
\pause
\vskip.2in
\begin{beamerboxesrounded}[upper=palette tertiary, shadow=true]{Lemma 5}
    The executives face the following trade off when choosing $\bta$: higher tariffs decrease the probability that the trade agreement will be broken, but they also decrease welfare when the agreement is in force.
\end{beamerboxesrounded}
\end{frame}

\begin{frame}{Trade Agreements}

\begin{beamerboxesrounded}[upper=palette tertiary, shadow=true]{Result 4}
  The executives maximize their welfare by either (a) trading off reductions in the probability that the agreement will be broken with reductions in welfare under the agreement or (b) raising tariffs sufficiently high to ensure that the trade agreement will remain in force.
\end{beamerboxesrounded}
\end{frame}

\begin{comment}
\begin{frame}{Comparing Government Structures}
\begin{changemargin}{-1cm}{-1cm}
\begin{table}
\centering
  \begin{tabular}{lrrrr}
		\toprule
			Structure & $\bta$ & Lobbying & $B(\bta)$ & $\expect[\bta]$ \\
			\midrule
			Unitary, No Agreement & ---- & 0.0017 & 0 & 0.129 \\
			Unitary, Agreement & 0.118 & 0.0027 & 0 & 0.118 \\
			\\
			Exec-led Agreement (.25) & 0.078 & 0.0007 & 0.48 & 0.102 \\
			Exec-led Agreement (.00) & 0.106 & 0.0000 & 0 & 0.106 \\
			& {\color{gray}(.105)} & {\color{gray}(.0025)} & {\color{gray}(1)} & {\color{gray}(.129)} \\
		\bottomrule
	\end{tabular}
	\end{table}
\end{changemargin}
\end{frame}

\begin{frame}{Comparative Systems}
\begin{beamerboxesrounded}[upper=palette tertiary, shadow=true]{Conjecture}
    Lobbying effort and tariffs are reduced by the presence of executive branches who form a trade agreement.
\end{beamerboxesrounded}

\pause
\vskip.2in
\begin{beamerboxesrounded}[upper=palette tertiary, shadow=true]{Corollary}
    Tariffs are lower in countries with separated powers than in those with unitary governments.
\end{beamerboxesrounded}
\end{frame}
\end{comment}

\begin{frame}[label=Uncertainty]
\frametitle{Uncertainty}
\begin{itemize}[<+->]
	\item Added intuition for lobbying incentives and tariff-setting behavior
  \item High tariffs, low but positive lobbying effort
  \item Possibility of trade disruptions / ratification failure in equilibrium
  \item Different behavior at varying levels of uncertainty
\end{itemize}
\end{frame}

\section{Conclusion}
\subsection{}
\begin{frame}{Future Work}
\begin{itemize}[<+->]
	\item Comparative statics on weighting function
	\item Empirical Tests
	\item Parallel models for countries with different government structures
	\item Asymmetric agreements
	\item Repeated model
		\begin{itemize}
			\item Implications for design of dispute settlement procedures
		\end{itemize}
\end{itemize}
\end{frame}

\begin{frame}{Conclusion}
Resolution of `Protection for Sale' Puzzle
\pause
\begin{itemize}[<+->]
	\item Non-unitary legislature: more realistic lobbying process
	\item Separation of Powers: tariffs set in shadow of lobbying
	\item Uncertainty: smoother results, additional intuition
\end{itemize}
\end{frame}

\begin{frame}{Conclusion}
\begin{itemize}[<+->]
	\item Social welfare maximizers ($a \approx \infty$) set high tariffs to achieve low lobbying expenditures (and low $\ga_{ML}$) 
	\item Executives use trade agreement to manipulate lobbying incentives
		\begin{itemize}
			\item Trade agreement as political commitment device
		\end{itemize}
	\item Firms lobby for trade war / ratification failure and achieve equilibrium trade disruptions in some cases
		\begin{itemize}
			\item Can examine the endogenous dynamics behind politically-driven failures to cooperate
		\end{itemize}
\end{itemize}

\end{frame}

\end{document}